Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report

### 2023-07-09 manuscript.tex
Updated version

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report},
  pdfauthor={Matthew H.C. Mak, Alice O'Hagan, Aidan J. Horner, M. Gareth Gaskell},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

For this question, the dependent variable was whether a critical lure is recalled or not (i.e., binary). There were two independent variables: Interval (Immediate vs.~Delay) and Test Time (9AM vs.~9PM), both of which were manipulated between-participants. In other words, the four groups were coded as in Table 1:

@@

We made use of 20 DRM wordlists (see Table 2), taken from Roediger et al. {[}\protect\hyperlink{ref-roediger2001a}{21}{]}. Each list contained 8 semantically related words, and as per the standard DRM paradigm, they were arranged in a descending order of associative strength to the critical lures. A participant studied all 20 lists. We note that the original DRM lists by Roediger et al. {[}\protect\hyperlink{ref-roediger2001a}{21}{]} were tailored for American participants, and two words (e.g., trash, Mississippi) were not immediately relatable to people in the UK. We, therefore, changed these words (e.g., trash → rubbish), as noted in Table 2.

@@

During presentation, words in each DRM list were presented visually,\footnote{Newbury and Monaghan [@newbury2019a] found no evidence in their meta-analysis that the modality of presentation modulated the effect of sleep. However, in a set of three experiments, Fenn et al. [@fenn2009a] used auditory presentation in one of them and visual in the other two. As far as we are aware, this is the only study in the ‘Sleep x DRM’ literature that had used both modalities in the same set of experiments. Sleep appeared to have a larger effect on false memory when visual (vs. auditory) presentation was used. Given this, we opted for visual presentation in the current experiment. } in a fixed order and arranged in descending associative strength to the unpresented critical lure (see Table 2 for order). Each list began with a fixation for 1 s, followed by the first word in a list. Each word was shown for 1 s, in a lowercase black font (Arial, size 26) on a white background, and separated by a 500-ms interstimulus interval. After presentation of the final word in a list was 5 s of blank screen. List order was randomised, and each list was seen once.

@@

Of the 535 participants who completed both the study and test phases, 47 were excluded for meeting one or more of the exclusion criteria. The full list of excluded participants and their respective reason for exclusion is available on OSF (see exclusion\_OSF.csv). Our final sample size comprised 488 participants, with 124 in each of the Immediate groups, and 120 in each of the Delay groups. Group characteristics are summarised in Table 3.

Prior to the confirmatory analyses, we first checked if the four groups were matched on their morningness/eveningness preference and degree of sleepiness at study/test (as indexed by the Stanford Sleepiness Scale). These are summarised in Table 3. We compared the four groups on each of the measures using one-way ANOVAs, which showed no significant differences (SSS at study: F = 0.21, p = .888; SSS at test: F = 1.63, p = .183; rMEQ: F = 0.97, p = .408). We also compared the Sleep and Wake groups on the number of intervening hours between study and test using a between-participant t-test, which revealed no significant difference {[}t(236.16) = 0.86, p = .389{]}. In sum, our four groups were well-matched on these potentially confounding factors.

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance \(\leq\) 2 (e.g., *cigerette → cigarette).\footnote{If a misspelt response has more than one nearest word, the response was considered as an intrusion.} Responses with added or dropped inflectional suffixes (i.e., -s, -ed, -ing, adjectival -er) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is pollution; if a participant recalled pollutions, the plural suffix was dropped; however, if a participant recalled pollutant, this was considered as an intrusion.
\# Results of Pre-registered Analyses

@@

We checked if our paradigm consistently elicited the well-established DRM effect across participants. Given free recall, the chance level of a critical lure being produced is 0. We submitted the number of critical lures produced by all participants (Range: 0 to 20) to a one-sample t-test, with the chance level being 0. It showed that participants were susceptible to false recall {[}t(487) = 26.96, p \textless{} .001{]}, providing evidence for the classic DRM false memory effect.

@@

A 2 \(\times\) 2 Poisson regression revealed significant effects of Interval (\(\beta\) = 0.372, SE = 0.054, z = 6.845, p \textless{} .001) and Test Time (\(\beta\) = 0.185, SE = 0.056, z = 3.299, p \textless{} .001), which were qualified by a significant interaction (\(\beta\) = 0.214, SE = 0.072, z = 2.96, p = .003). Given this, we tested the simple effects of Test Time within the Immediate and Delay groups using the emmeans package {[}\protect\hyperlink{ref-lenth2021a}{25}{]}. Within the Immediate groups, the evening participants (M = 5.62; SD = 6.63) produced more intrusions than the morning participants (M = 4.67, SD = 5.97) (z = -3.299, p = .001). Likewise, in the Delay groups, the Wake participants, who completed free recall in the evening (M = 10.1, SD = 13.65), produced more intrusions than the Sleep participants, who completed recall in the morning (M = 6.78; SD = 8.71) (z = -8.808, p = \textless.001). Together, our data indicate that participants who attempted free recall in the evening (vs.~morning) were more prone to intrusions, and this effect was greater in the Delay than in the Immediate groups. These unexpected findings prompted us to explore whether the number of total responses (i.e., studied + lures + intrusions) differed between morning and evening test time. Interestingly, this exploratory analysis (see section 13.1) showed no effect of Test Time. Together, these suggest that attempting free recall in the evening led to a selective increase in intrusions, but not necessarily a global increase in output bias. Finally, given that Test Time had a significant effect on intrusions, we followed our pre-registered analysis plan by adding However, neither Test Time nor their interaction is expected to be significant. If in the unlikely event that these are revealed to be significant, the number of intrusions as a numeric covariate in the 2 \(\times\) 2 mixed-effects models below.

@@

A generalised linear mixed-effect model (GLMM) was fitted to the critical lure data on the item level (N of observations = 488 participants \(\times\) 20 critical lures).\footnote{The use of GLMM is a clear departure from prior ‘Sleep $\times$ DRM’ studies, the majority of which addressed the same research question using an independent t-test or ANOVA (e.g., Diekelmann et al., 2010; Monaghan et al., 2017; Payne et al., 2009). We explained in Appendix C why these statistical tests are usually not appropriate in the context of DRM recall and why GLMM are more advantageous.} The dependent variable was binary: whether a critical lure was recalled or not (1 vs.~0). The fixed effects were the number of intrusions a participant produced, Interval (Immediate vs.~Delay), Test Time (AM vs.~PM), and an Interval by Test Time interaction. Interval and Test Time were coded using sum contrasts {[}\protect\hyperlink{ref-barr2019a}{26}{]}. The random-effect structure was determined by the ``buildmer'' package {[}\protect\hyperlink{ref-voeten2021a}{27}{]}, which automatically found the maximal model that was capable of converging using backward elimination (with the ``bobyqa'' optimiser). This means that model selection started from the maximal model, as justified by the experimental design {[}\protect\hyperlink{ref-barr2013a}{28}{]}. The model we reported and based our interpretation on was the most maximal model that was capable of converging (see the upper half of Table 4 for the final random-effect structure and model output).

The number of intrusions had a significant effect on lure recall, such that participants who produced more intrusions tended to recall more critical lures (z = 5.487, p \textless{} .001). There were no main effects of Interval or Test Time (zs \textless{} 0.95, ps \textgreater{} .34), but there was a significant Interval by Test Time interaction (z = -2.046, p = .041). Following our pre-registered analysis plan, we proceeded to test the simple effects of Test Time within the Immediate and Delay groups, using the emmeans package {[}\protect\hyperlink{ref-lenth2021a}{25}{]} in R. Among the Immediate groups, there was no significant difference in lure recall between the AM-control and PM-control participants (\(\beta\) = -0.098, SE = 0.114, z = -0.859, p = .390). However, among the Delay groups, there was a significant difference (\(\beta\) = 0.239, SE = 0.119, z = 2.00, p = .045) such that the Sleep participants (M = 2.73, SD = 2.30) reported more critical lures than the Wake participants (M = 2.51, SD = 2.32).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report},
  pdfauthor={Matthew H.C. Mak, Alice O'Hagan, Aidan J. Horner, M. Gareth Gaskell},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

For this question, the dependent variable was whether a critical lure is recalled or not (i.e., binary). There were two independent variables: Interval (Immediate vs.~Delay) and Test Time (9AM vs.~9PM), both of which were manipulated between-participants. In other words, the four groups were coded as in Table \ref{tab:table1}:

\begin{table}[H]

\caption{\label{tab:table1}How the four groups were coded using Interval and Test Time.}
\centering
\begin{tabular}[t]{ccccc}
\toprule
\textbf{Groups} & \textbf{ } & \textbf{Interval} & \textbf{ } & \textbf{Test Time}\\
\midrule
AM-control & = & Immediate & + & 9AM\\
PM-control & = & Immediate & + & 9PM\\
Sleep & = & Delay & + & 9AM\\
Wake & = & Delay & + & 9PM\\
\bottomrule
\end{tabular}
\end{table}

@@

We made use of 20 DRM wordlists (see Table \ref{tab:table2}), taken from Roediger et al. {[}\protect\hyperlink{ref-roediger2001a}{21}{]}. Each list contained 8 semantically related words, and as per the standard DRM paradigm, they were arranged in a descending order of associative strength to the critical lures. A participant studied all 20 lists. We note that the original DRM lists by Roediger et al. {[}\protect\hyperlink{ref-roediger2001a}{21}{]} were tailored for American participants, and two words (e.g., trash, Mississippi) were not immediately relatable to people in the UK. We, therefore, changed these words (e.g., trash → rubbish), as noted in Table \ref{tab:table2}.

@@

During presentation, words in each DRM list were presented visually,\footnote{Newbury and Monaghan [@newbury2019a] found no evidence in their meta-analysis that the modality of presentation modulated the effect of sleep. However, in a set of three experiments, Fenn et al. [@fenn2009a] used auditory presentation in one of them and visual in the other two. As far as we are aware, this is the only study in the ‘Sleep $\times$ DRM’ literature that had used both modalities in the same set of experiments. Sleep appeared to have a larger effect on false memory when visual (vs. auditory) presentation was used. Given this, we opted for visual presentation in the current experiment.} in a fixed order and arranged in descending associative strength to the unpresented critical lure (see Table \ref{tab:table2} for order). Each list began with a fixation for 1 s, followed by the first word in a list. Each word was shown for 1 s, in a lowercase black font (Arial, size 26) on a white background, and separated by a 500-ms interstimulus interval. After presentation of the final word in a list was 5 s of blank screen. List order was randomised, and each list was seen once.

\begin{table}[H]

\caption{\label{tab:table2}The 20 DRM wordlists to be used.}
\centering
\begin{tabular}[t]{lll}
\toprule
\textbf{Critical lure} & \textbf{False recall probability} & \textbf{List items (arranged in the}\\
\textbf{of each list} & \textbf{(Roediger et al. 2001)} & \textbf{order of presentation in study)}\\
\midrule
\em{Window} & \em{65} & \em{door, glass, pane, shade, ledge, sill, house, open}\\
\em{Sleep} & \em{61} & \em{bed, rest, awake, tired, dream, wake, snooze, blanket}\\
\em{Doctor} & \em{60} & \em{nurse, sick, lawyer, medicine, health, hospital, dentist, physician}\\
\addlinespace
\em{Smell} & \em{60} & \em{nose, breathe, sniff, aroma, hear, see, nostril, whiff}\\
\em{Chair} & \em{54} & \em{table, sit, legs, seat, couch, desk, recliner, sofa}\\
\em{Smoke} & \em{54} & \em{cigarette, puff, blaze, billows, pollution, ashes, cigar, chimney}\\
\em{Sweet} & \em{54} & \em{sour, candy, sugar, bitter, good, taste, tooth, nice}\\
\em{Rough} & \em{53} & \em{smooth, bumpy, road, tough, sandpaper, jagged, ready, coarse}\\
\addlinespace
\em{Needle} & \em{52} & \em{thread, pin, eye, sewing, sharp, point, prick, thimble}\\
\em{Rubbish} & \em{49} & \em{garbage, waste, can, refuse, sewage, bag, junk, trash (Note 1)}\\
\em{Anger} & \em{49} & \em{mad, fear, hate, rage, temper, fury, ire, wrath}\\
\em{Soft} & \em{46} & \em{hard, light, pillow, plush, loud, cotton, fur, touch}\\
\em{City} & \em{46} & \em{town, crowded, state, capital, streets, subway, country, New York}\\
\addlinespace
\em{Cup} & \em{45} & \em{mug, saucer, tea, measuring, coaster, lid, handle, coffee}\\
\em{Cold} & \em{44} & \em{hot, snow, warm, winter, ice, wet, frigid, chilly}\\
\em{Mountain} & \em{42} & \em{hill, valley, climb, summit, top, molehill, peak, plain}\\
\em{Slow} & \em{42} & \em{fast, lethargic, stop, listless, snail, cautious, delay, traffic}\\
\em{River} & \em{42} & \em{water, stream, lake, Thames (Note 2), boat, tide, swim, flow}\\
\addlinespace
\em{Spider} & \em{37} & \em{web, insect, bug, fright, fly, arachnid, crawl, tarantula}\\
\em{Foot} & \em{35} & \em{shoe, hand, toe, kick, sandals, soccer, yard, walk}\\
\bottomrule
\multicolumn{3}{l}{\textsuperscript{} Note 1. In Roediger et al. (2001), the critical lure for this list was trash, with rubbish being one of the list}\\
\multicolumn{3}{l}{items. We used rubbish as the critical lure and trash as a list item because the former is the preferred term in}\\
\multicolumn{3}{l}{British English.}\\
\multicolumn{3}{l}{\textsuperscript{} Note 2. The original word in Roediger et al. was Mississippi. We replaced it with Thames.}\\
\end{tabular}
\end{table}

@@

Of the 534 participants who completed both the study and test phases, 46 were excluded for meeting one or more of the exclusion criteria. The full list of excluded participants and their respective reason for exclusion is available on OSF (see exclusion\_OSF.csv). Our final sample size comprised 488 participants, with 124 in each of the Immediate groups, and 120 in each of the Delay groups. Group characteristics are summarised in Table \ref{tab:table3}.

\begin{table}[H]

\caption{\label{tab:table3}Group characteristics.}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Characteristics & Immediate & Immediate & Sleep (aka & Wake (aka\\
 & -AM & -PM & Delay-AM) & Delay-PM)\\
\midrule
N before exclusion & 130 & 127 & 135 & 143\\
N after exclusion & 124 & 124 & 120 & 120\\
Mean age (SD) & 22.24 (2.18) & 22.34 (2.03) & 22.18 (1.93) & 22.25 (1.93)\\
\addlinespace
Gender (Female:Male:Other) & 64 : 54 : 2 & 77 : 46 : 1 & 62 : 57 : 1 & 58 : 61 : 1\\
\% participants identified as ethnically white & 78.2\% & 73.4\% & 81.7\% & 80\%\\
Mean SSS rating at study (SD) & 2.58 (0.98) & 2.64 (1.12) & 2.66 (0.96) & 2.58 (0.98)\\
Mean SSS rating at test (SD) & 2.73 (1.04) & 2.95 (1.29) & 2.63 (1.21) & 2.67 (1.18)\\
Mean rMEQ score (SD) & 15.89 (1.67) & 15.59 (1.91) & 15.72 (1.83) & 15.53 (1.99)\\
\addlinespace
Mean N of intervening hr between study and test (SD) & NA & NA & 12.22 (0.74) & 12.14 (0.81)\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Notes. (1) SSS stands for Stanford Sleepiness Scale and ranges from 1 to 6, with higher values indicating}\\
\multicolumn{5}{l}{greater sleepiness. (2) rMEQ stands for reduced Morningness/Eveningness Questionnaire; it ranges from 1}\\
\multicolumn{5}{l}{to 25, with higher values indicating greater morningness preference.}\\
\end{tabular}
\end{table}

Prior to the confirmatory analyses, we first checked if the four groups were matched on their morningness/eveningness preference and degree of sleepiness at study/test (as indexed by the Stanford Sleepiness Scale). These are summarised in Table \ref{tab:table3}. We compared the four groups on each of the measures using one-way ANOVAs, which showed no significant differences (SSS at study: F = 0.21, p = .888; SSS at test: F = 1.63, p = .183; rMEQ: F = 0.97, p = .408). We also compared the Sleep and Wake groups on the number of intervening hours between study and test using a between-participant t-test, which revealed no significant difference {[}t(236.16) = 0.86, p = .389{]}. In sum, our four groups were well-matched on these potentially confounding factors.

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance \(\leq\) 2 (e.g., *cigerette → cigarette).\footnote{If a misspelt response has more than one nearest word, the response was considered as an intrusion.} Responses with added or dropped inflectional suffixes (i.e., -s, -ed, -ing, adjectival -er) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is pollution; if a participant recalled pollutions, the plural suffix was dropped; however, if a participant recalled pollutant, this was considered as an intrusion.

\hypertarget{results-of-pre-registered-analyses}{%
\section{Results of Pre-registered Analyses}\label{results-of-pre-registered-analyses}}

@@

We checked if our paradigm consistently elicited the well-established DRM effect across participants. Given free recall, the chance level of a critical lure being produced is 0. We submitted the number of critical lures produced by all participants (Range: 0 to 20) to a one-sample t-test, with the chance level being 0. It showed that participants were susceptible to false recall {[}t(487) = 26.96, \textless0.001{]}, providing evidence for the classic DRM false memory effect.

@@

A 2 \(\times\) 2 Poisson regression revealed significant effects of Interval (\(\beta\) = 0.372, SE = 0.054, z = 6.845, p \textless{} \textless0.001) and Test Time (\(\beta\) = 0.185, SE = 0.056, z = 3.299, p \textless{} \textless0.001), which were qualified by a significant interaction (\(\beta\) = 0.214, SE = 0.072, z = 2.96, p = 0.003). Given this, we tested the simple effects of Test Time within the Immediate and Delay groups using the emmeans package {[}\protect\hyperlink{ref-lenth2021a}{25}{]}. Within the Immediate groups, the evening participants (M = 5.62; SD = 6.63) produced more intrusions than the morning participants (M = 4.67; SD = 5.97) (z = -3.299, p = \textless0.001). Likewise, in the Delay groups, the Wake participants, who completed free recall in the evening (M = 10.1; SD = 13.65), produced more intrusions than the Sleep participants, who completed recall in the morning (M = 6.78; SD = 8.71) (z = -8.808, p = \textless0.001). Together, our data indicate that participants who attempted free recall in the evening (vs.~morning) were more prone to intrusions, and this effect was greater in the Delay than in the Immediate groups. These unexpected findings prompted us to explore whether the number of total responses (i.e., studied + lures + intrusions) differed between morning and evening test time. Interestingly, this exploratory analysis (see section 13.1) showed no effect of Test Time. Together, these suggest that attempting free recall in the evening led to a selective increase in intrusions, but not necessarily a global increase in output bias. Finally, given that Test Time had a significant effect on intrusions, we followed our pre-registered analysis plan by adding However, neither Test Time nor their interaction is expected to be significant. If in the unlikely event that these are revealed to be significant, the number of intrusions as a numeric covariate in the 2 \(\times\) 2 mixed-effects models below.

@@

A generalised linear mixed-effect model (GLMM) was fitted to the critical lure data on the item level (N of observations = 488` participants \(\times\) 20 critical lures).\footnote{The use of GLMM is a clear departure from prior ‘Sleep $\times$ DRM’ studies, the majority of which addressed the same research question using an independent t-test or ANOVA (e.g., Diekelmann et al., 2010; Monaghan et al., 2017; Payne et al., 2009). We explained in Appendix C why these statistical tests are usually not appropriate in the context of DRM recall and why GLMM are more advantageous.} The dependent variable was binary: whether a critical lure was recalled or not (1 vs.~0). The fixed effects were the number of intrusions a participant produced, Interval (Immediate vs.~Delay), Test Time (AM vs.~PM), and an Interval by Test Time interaction. Interval and Test Time were coded using sum contrasts {[}\protect\hyperlink{ref-barr2019a}{26}{]}. The random-effect structure was determined by the ``buildmer'' package {[}\protect\hyperlink{ref-voeten2021a}{27}{]}, which automatically found the maximal model that was capable of converging using backward elimination (with the ``bobyqa'' optimiser). This means that model selection started from the maximal model, as justified by the experimental design {[}\protect\hyperlink{ref-barr2013a}{28}{]}. The model we reported and based our interpretation on was the most maximal model that was capable of converging (see the upper half of Table \ref{tab:table4} for the final random-effect structure and model output).

\begin{table}[H]

\caption{\label{tab:table4}Outputs from confirmatory generalised mixed-effect models examining the effects of Intrusions, Interval, and Test Time in false and veridical recall.}
\centering
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{5}{c}{\textbf{False (lure) recall}} \\
\cmidrule(l{3pt}r{3pt}){1-5}
\multicolumn{5}{l}{Random-effect structure: (Intrusions | Participant.ID) + (1 | Lure)} \\
\cmidrule(l{3pt}r{3pt}){1-5}
\em{} & \em{Estimate} & \em{SE} & \em{z} & \em{p}\\
Intercept & -2.328 & 0.11 & -21.254 & <.001*\\
Intrusions & 0.034 & 0.006 & 5.487 & <.001*\\
Interval (Immediate vs. Delay) & 0.039 & 0.042 & 0.943 & .346\\
Test Time (AM vs. PM) & 0.035 & 0.041 & 0.85 & .395\\
\addlinespace
Interval x Test Time & -0.084 & 0.041 & -2.046 & .041*\\
\bottomrule
\end{tabular}
\end{table}
